---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.align = "center",
  fig.path = "man/figures/README-")
```

# qoa

QOA is fast. It decodes audio 3x faster than Ogg-Vorbis, while offering
better quality and compression (278 kbits/s for 44khz stereo) than
ADPCM.

<https://qoaformat.org/>

This **R**-package is a wrapper to handle QOA files with **R**. The
package is completely written in base R with no dependencies and pure
C-code. To handle the raw audio data the R-package `tuneR`is highly recommended.

See this blogpost for an introduction <https://phoboslab.org/log/2023/02/qoa-time-domain-audio-compression>

<!-- badges: start -->
[![Project Status: Active -- The project has reached a stable, usable
state and is being actively
developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

<!-- badges: end -->

## Installation

You can install the development version from GitHub with the following
command:

```{r, eval = FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("JohannesFriedrich/qoa4R")
```

## Usage

There are just two main functions: `readQOA()` and `writeQOA()`.

-   `readQOA()`: Takes an qoa-format audio file and decodes it into it´s
    raw audio data. Return value is a list with decoded values
    represented as matrix, used channels, sample rate and samples.
-   `writeQOA()`: Takes an matrix with every channel in new column and
    the sample rate as input and and encodes it into an qoa audio file.

### `readQOA()`

Let´s read in an QOA-file delivered with this package:

```{r}
library(qoa)
qoa_file <- system.file("extdata", "58_guitar_sarasate_stereo.qoa", package = "qoa")
qoa_data <- readQOA(qoa_file)
```

`qoa_data` is a list with the following elements:
* `data`: a matrix with all the samples. The dimension of the matrix is samples x channels
* `channels`: Number of channels 
* `samples`: Number of samples per channel
* `samplerate`: Sample rate of the file

It is possible to write these samples and information with the `tuneR` package back to an wave file:

```{r, eval = FALSE}
tuneR::writeWave(
  WaveMC(
    data = qoa_data$data, 
    samp.rate = qoa_data$samplerate, 
    bit = 16), 
  "wave_from_qoa.wav", 
  extensible = FALSE)

```


### `writeQOA()`

With this function it is possible to write an qoa-file. The input arguments are:

-   samples: a matrix with dimensions samples x channels 
-   samplerate: samplerate of the audio file
-   target: Either name of the file to write, a binary connection or a
    raw vector indicating that the output should be a raw vector (the
    hex-interpretation of the qoa-file).

If no second argument is given, the returned value is the
hex-interpretation of the image in the QOI-file format.

```{r, eval=FALSE}
wav_original <- tuneR::readWave("wave_from_qoa.wav", toWaveMC = TRUE)
writeQOA(samples = wav_original@.Data, 
         samplerate = wav_original@samp.rate)
```

If an second argument is given as character the image is saved to this
name:

```{r, eval = FALSE}
writeQOA(samples = wav_original@.Data, 
         samplerate = wav_original@samp.rate, 
         target = "test.qoa")
```

If the second argument is of type `connection` the hex interpretation of
the audio file will be send to this connection, e.g. via sockets or FTP.

```{r, eval = FALSE}
file <- file("file.qoa", "wb")
writeQOA(samples = wav_original@.Data, 
         samplerate = wav_original@samp.rate, 
         target = "test.qoa")
close(file)
```

## Acknowlegment

This package would not exist without the following
persons/homepages/tutorial/...:

-   [Phoboslab - original
    specification](https://github.com/phoboslab/qoa)
-   [Example C-code
    R-package](https://github.com/coolbutuseless/simplecall)
-   [R
    extensions](https://cran.r-project.org/doc/manuals/r-release/R-exts.html)
-   [Hadley´s R-Internals](https://github.com/hadley/r-internals)
